---
title: "Key Skills in <br/> Statistics, <br/> Data Science, <br/> and Statistical Programming  <br/> for Resource Managers"
subtitle: "(Conducted under EMF 115 1.0 Key Skills for Resource Managers)"
author: "Dr. Thiyanga S. Talagala "
format:
  revealjs:
    css:
        - "custom.css"
    width: 1600
    height: 900
    margin: 0.1
    theme: simple
    transition: slide
    background-transition: fade
    slide-number: true
    show-slide-number: all
    title-slide-attributes: 
      data-background-color: "#016450"
      data-background-image: none
---

# Recap

- Why Statistics is needed for managers

- Statistician/ Data Scientist: What they do and How they work


- Primary Data vs Secondary Data


- Primary Data Collection Methods

- Some tools for data analysis

# Data Visualisation

---

```{r, echo=TRUE}
library(palmerpenguins)
data("penguins")
library(dplyr)
penguins <- penguins |>
  select(1, 2, 5, 6)
as.data.frame(penguins) |> knitr::kable()
```
---

## Step 1: Identify qualitative and quantitative variables

```{r}
as.data.frame(penguins) |> knitr::kable()

```

---

# Step 2: Qualitative Data Analysis: Univariate Analysis

---

### Composition of the sample by species

```{r, echo=TRUE}
library(ggplot2)
ggplot(penguins, aes(x=species)) + geom_bar()
```

---

### Composition of the sample by species

```{r, echo=TRUE}
library(ggplot2)
p1 <- ggplot(penguins, aes(x=species, fill=species)) + geom_bar()
p1
```

---

### Selecting colours

Go to [https://colorbrewer2.org/](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3)

Sri Lanka Nature Inspired Colour Palettes [https://github.com/thiyangt/nic](https://github.com/thiyangt/nic)

---
## Pie chart

```{r}
#| warning: false
#| message: false
## Pie chart
library(magrittr)
library(tidyverse)
# Remove rows with missing species data
pf <- penguins %>%
  drop_na(species)

# Create the frequency table with counts and percentages
frequency_table <- pf %>%
  count(species) %>%
  mutate(percentage = n / sum(n) * 100)

# Plot pie chart
pie_chart <- ggplot(frequency_table, aes(x = "", y = n, fill = species)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start=0) +
  labs(title = "Composition of the Sample by Species") +
  theme_void() +
  theme(legend.position = "bottom")
pie_chart
```



# Which is the best?

##

:::: {.columns}
::: {.column width="50%"}
```{r}
p1
```
:::

::: {.column width="50%"}
```{r}
pie_chart
```
:::
::::



# Step 3: Qualitative Data Analysis: Bivariate Analysis


#
  
```{r, echo=TRUE}
library(ggplot2)
ggplot(penguins, aes(x=species, fill=island)) + geom_bar() 
```


#
  
```{r, echo=TRUE}
library(ggplot2)
ggplot(penguins, aes(x=species, fill=island)) + geom_bar()  + scale_fill_brewer(palette="Set2")
```


#
  
```{r, echo=TRUE}
library(ggplot2)
ggplot(penguins, aes(x = species, fill = island)) + 
  geom_bar(position = position_dodge2(preserve = "single", width = 0.9), width = 0.8) + 
  scale_fill_brewer(palette = "Set2")

```


# Step 4: Quantitative Data Analysis: Univariate Analysis

## Histogram

```{r}
ggplot(penguins, aes(x = body_mass_g)) + geom_histogram(col="white")

```

## Histogram

```{r}
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram(col="white")

```

# Step 5: Quantitative Data Analysis: Bivariate Analysis

## Scatterplot

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y=body_mass_g)) + geom_point()
```